\bf A simple algorithm for computing approximate Nash equilibria in two-person zero-sum games with imcomplete information

### 2015-06-07 nips2015.tex
more

\begin{abstract}
We present a simple primal-dual algorithm
%with low costs per iteration,
for computing approximate Nash equilibria in two-person zero-sum
sequential games with imcomplete information and perfect recall (like
Texas Hold'em Poker).
Our algorithm only performs basic iterations
(i.e iterations involving no
% only involving basic operations like matvec, clipping, etc., and no
calls to external first-order oracles, etc.) and is applicable to a
broad class of two-person zero-sum
games including simultaneous games and sequential games with
imcomplete information and perfect recall. The applicability to the
latter kind of games is thanks to the sequence-form representation von
Stengel's \cite{von1996efficient}.
%% Unlike the EGT (Excessive Gap Technique) recently used in
%% \cite{hoda2010smoothing, gilpinfirst}, our algorithm has a cheap
%% and constant cost per iteration.
As an application, we apply our algorithm to solve Kuhn 3-card Poker.
%% We conclude by extending our results to tackle Nash equilibrium problems with even more general strategy profiles, where the nonnegativity constraints on realization plans are replaced with more general convex sets.
\end{abstract}

@@

\paragraph{\textbf{Statement of the problem.}}
The sequence-form for two-person zero-sum games was introduced in
\cite{koller1992complexity}, and the theory was further developed in
\cite{von1996efficient,vonequilibrium}, where it was established that
for sequential two-person zero-sum games with imcomplete information
and perfect recall, there exist sparse matrices $A \in \mathbb{R}^{n_1
  \times n_2}$, $E_1 \in \mathbb{R}^{l_1 \times n_1}$, $E_2 \in
\mathbb{R}^{l_2 \times n_2}$, and vectors $e_1 \in \mathbb{R}^{l_1},
e_2 \in \mathbb{R}^{l_2}$ such that $n_1$, $n_2$, $l_1$, and $l_2$ are
all linear in the size of the game tree (number of states in the game)
and such that Nash equilibria correspond to pairs $(x, y)$ of
\textit{realization plans} which solve the primal LCP
\begin{equation}
  \begin{aligned}
     \text{minimize }e_1^Tp, \hspace{.5em}\text{ subject to }(y,p) \in
     \mathbb{R}^{n_2} \times \mathbb{R}^{l_1}, -Ay +
    E_1^Tp \geq 0, E_2y = e_2, y \ge 0.
  \end{aligned}
  \label{eq:primal_pb}
\end{equation}

@@

\section{The algorithm}
\label{sec:algo}

%% According to Theorem \ref{thm:pd}, solving the Nash equilibrium LCPs
%% \eqref{eq:primal_pb}, \eqref{eq:dual_pb} is equivalent to solving the
%% saddle-point problem \eqref{eq:unconstrained_pb}. An attractive method
%% for such a saddle-point problem would be a primal-dual scheme, like
%% that developed in \cite[Algorithm 39]{chambolle2010}. 
%% Usually, to use a primal-dual algorithm, one needs a measure of the
%% ``quality'' of the current iterates as regards being the solution of
%% the saddle-point problem. A natural measure of this fittness is the
%% primal-dual gap, which measures defection from strong-duality.
%% However, the primal-dual gap in \eqref{eq:unconstrained_pb} can be
%% unbounded even at near feasible points.  is well-suited for

Consider the following algorithm
\begin{algorithm}
\caption{Primal-dual algorithm for computing approximate Nash
  Equilbria in two-person zero-sum games with imcomplete information
  and perfect recall}
\label{Tab:algo}
\begin{algorithmic}[1]
\Require $\rho >0$, $\epsilon > 0$; $(y^{(0)},p^{(0)},x^{(0)},q^{(0)})
\in \mathbb{R}^{n_2}
  \times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1} \times
  \mathbb{R}^{l_2}$.
\Ensure A Nash $(\rho, \epsilon)$-equilibrium
$(\tilde{y},\tilde{p},\tilde{x},\tilde{q}) \in S_1 \times S_2$ for
the GSP \eqref{eq:unconstrained_pb}.
\State  $\lambda \leftarrow 1/\|K\|$,
%% $y^{(0)} \leftarrow 0_{n_2}$,
%% $p^{(0)} \leftarrow 0_{l_1}$, $x^{(0)} \leftarrow 0_{n_1}$, $q^{(0)}
%% \leftarrow 0_{l_2}$,
$k \leftarrow 0$
\While{$\|\tilde{\epsilon}^{a}_k\| \ge \rho$ or $\|\tilde{v}^{a}_k\| \ge \epsilon$}
\State $y^{(k + 1)} \leftarrow (y^{(k)} - \lambda (A^Tx^{(k)} +
E_2^Tq^{(k)}))_+$
\State  \hspace{.5em}$p^{(k+1)} \leftarrow p^{(k)} -
\lambda(e_1-E_1x^{(k)})$
\State $x^{(k + 1)} \leftarrow (x^{(k)} + \lambda (Ay^{(k+1)} -
E_1^Tp^{(k+1)}))_+$
\State  \hspace{.5em}$q^{(k+1)} \leftarrow q^{(k)} +
\lambda(E_2y^{(k+1)}-e_2)$
\State $y^{(k+1)} \leftarrow y^{(k+1)} - \lambda
(A^T(x^{(k+1)}-x^{(k)}) + E_2^T(q^{(k+1)}-q^{(k)}))$
\State $p^{(k+1)} \leftarrow p^{(k+1)} - \lambda E_1(x^{(k+1)}-x^{(k)})$
\State $k \leftarrow k + 1$
\EndWhile
\end{algorithmic}
\end{algorithm}

>>>>

\begin{abstract}
We present a simple primal-dual algorithm for computing approximate
Nash equilibria in two-person zero-sum sequential games with
imcomplete information and perfect recall (like Texas Hold'em
Poker). Our algorithm only performs basic iterations (i.e iterations
involving no calls to external first-order oracles, etc.) and is
applicable to a broad class of two-person zero-sum games including
simultaneous games and sequential games with imcomplete information
and perfect recall. The applicability to the latter kind of games is
thanks to the sequence-form representation
\cite{koller1992complexity}. As an application, we apply our algorithm
to solve Kuhn 3-card Poker.
\end{abstract}

@@

\paragraph{\textbf{Statement of the problem.}}
The sequence-form for two-person zero-sum games was introduced in
\cite{koller1992complexity}, and the theory was further developed in
\cite{koller1994fast,von1996efficient,vonequilibrium}, where it was
established that for sequential two-person zero-sum games with
imcomplete information and perfect recall, there exist sparse matrices
$A \in \mathbb{R}^{n_1
  \times n_2}$, $E_1 \in \mathbb{R}^{l_1 \times n_1}$, $E_2 \in
\mathbb{R}^{l_2 \times n_2}$, and vectors $e_1 \in \mathbb{R}^{l_1},
e_2 \in \mathbb{R}^{l_2}$ such that $n_1$, $n_2$, $l_1$, and $l_2$ are
all linear in the size of the game tree (number of states in the game)
and such that Nash equilibria correspond to pairs $(x, y)$ of
\textit{realization plans} which solve the primal LCP
\begin{equation}
  \begin{aligned}
     \text{minimize }e_1^Tp, \hspace{.5em}\text{ subject to }(y,p) \in
     \mathbb{R}^{n_2} \times \mathbb{R}^{l_1}, -Ay +
    E_1^Tp \geq 0, E_2y = e_2, y \ge 0.
  \end{aligned}
  \label{eq:primal_pb}
\end{equation}

@@

\section{The algorithm}
\label{sec:algo}
\textbf{XXX TODO: Define the ergodic variables $\tilde{v}^a$ and
  $\tilde{\epsilon}^a$, ...}

Consider the following algorithm for solving the GSP \eqref{eq:unconstrained_pb}
\begin{algorithm}
\caption{Primal-dual algorithm for computing approximate Nash
  Equilbria in two-person zero-sum games with imcomplete information
  and perfect recall}
\label{Tab:algo}
\begin{algorithmic}[1]
\Require $\rho >0$, $\epsilon > 0$; $(y^{(0)},p^{(0)},x^{(0)},q^{(0)})
\in \mathbb{R}^{n_2}
  \times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1} \times
  \mathbb{R}^{l_2}$.
\Ensure A Nash $(\rho, \epsilon)$-equilibrium
$(\tilde{y},\tilde{p},\tilde{x},\tilde{q}) \in S_1 \times S_2$ for
the GSP \eqref{eq:unconstrained_pb}.
\State  $\lambda \leftarrow 1/\|K\|$,
%% $y^{(0)} \leftarrow 0_{n_2}$,
%% $p^{(0)} \leftarrow 0_{l_1}$, $x^{(0)} \leftarrow 0_{n_1}$, $q^{(0)}
%% \leftarrow 0_{l_2}$,
$k \leftarrow 0$
\While{$\|\tilde{\epsilon}^{a}_k\| \ge \rho$ or $\|\tilde{v}^{a}_k\| \ge \epsilon$}
\State $y^{(k + 1)} \leftarrow (y^{(k)} - \lambda (A^Tx^{(k)} +
E_2^Tq^{(k)}))_+$
\State  \hspace{.5em}$p^{(k+1)} \leftarrow p^{(k)} -
\lambda(e_1-E_1x^{(k)})$
\State $x^{(k + 1)} \leftarrow (x^{(k)} + \lambda (Ay^{(k+1)} -
E_1^Tp^{(k+1)}))_+$
\State  \hspace{.5em}$q^{(k+1)} \leftarrow q^{(k)} +
\lambda(E_2y^{(k+1)}-e_2)$
\State $y^{(k+1)} \leftarrow y^{(k+1)} - \lambda
(A^T(x^{(k+1)}-x^{(k)}) + E_2^T(q^{(k+1)}-q^{(k)}))$
\State $p^{(k+1)} \leftarrow p^{(k+1)} - \lambda E_1(x^{(k+1)}-x^{(k)})$
\State $k \leftarrow k + 1$
\EndWhile
\end{algorithmic}
\end{algorithm}
